### 2025-04-05 decision-making.qmd
20250405 - clarifications

As demonstrated above, when the individual predictions are centered on the truth, the bracketing rate is very high (in this case `{r} petersenlab::apa(bracketingCentered_bracketingRate, decimals = 2)`%), and the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).

@@

As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case `{r} petersenlab::apa(bracketingNotCentered1_bracketingRate, decimals = 2)`%) with no strong outliers, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).

@@

As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).

@@

As demonstrated above, when the individual predictions bracket but are not centered on the truth and the bracketing rate is low (in this case `{r} petersenlab::apa(bracketingnotCentered3_bracketingRate, decimals = 2)`%), the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered3_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered3_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered3_individual`).

>>>>

As demonstrated above, when the individual predictions are centered on the truth, the bracketing rate is very high (in this case, `{r} petersenlab::apa(bracketingCentered_bracketingRate, decimals = 2)`%), and the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).

@@

As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingNotCentered1_bracketingRate, decimals = 2)`%) with no strong outliers, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).

@@

As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).

@@

As demonstrated above, when the individual predictions bracket but are not centered on the truth and the bracketing rate is low (in this case, `{r} petersenlab::apa(bracketingnotCentered3_bracketingRate, decimals = 2)`%), the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered3_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered3_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered3_individual`).

### 2025-04-05 evaluating-prediction-accuracy.qmd
20250405 - clarifications

Mean absolute error (MAE) is the mean of the absolute value of differences between the predicted and observed values across individuals, that is, the mean of the absolute value of errors.
Smaller MAE values (closer to zero) reflect greater accuracy.
MAE is preferred over [root mean squared error](#sec-rootMeanSquaredError) (RMSE) when you want to give equal weight to all errors and when the outliers have considerable impact.
The formula for MAE is in @eq-meanAbsoluteError:

@@

Median absolute error (MdAE) is the median of the absolute value of differences between the predicted and observed values across individuals, that is, the median of the absolute value of errors.
Smaller MdAE values (closer to zero) reflect greater accuracy.
MdAE indicates the accuracy of the "typical" prediction (i.e., the prediction that is at the 50th percentile in terms of accuracy).
The formula for MdAE is in @eq-medianAbsoluteError:

>>>>

Mean absolute error (MAE) is the mean of the absolute value of differences between the predicted and observed values across individuals.
That is, MAE is the mean of the absolute value of errors (i.e., the mean distance between predicted and observed values).
Smaller MAE values (closer to zero) reflect greater accuracy.
MAE is preferred over [root mean squared error](#sec-rootMeanSquaredError) (RMSE) when you want to give equal weight to all errors and when the outliers have considerable impact.
The formula for MAE is in @eq-meanAbsoluteError:

@@

Median absolute error (MdAE) is the median of the absolute value of differences between the predicted and observed values across individuals.
That is, MdAE is the median of the absolute value of errors (i.e., the median distance between predicted and observed values).
Smaller MdAE values (closer to zero) reflect greater accuracy.
MdAE indicates the accuracy of the "typical" prediction (i.e., the prediction that is at the 50th percentile in terms of accuracy).
The formula for MdAE is in @eq-medianAbsoluteError:

### 2025-04-05 machine-learning.qmd
20250405 - clarifications

- supervised learning
    - continuous outcome (i.e., regression)
      - linear regression
      - lasso regression
      - ridge regression
      - elastic net regression
    - categorical outcome (i.e., classification)
      - logistic regression
      - support vector machine
      - random forest
      - extreme gradient boosting
- unsupervised learning
    - clustering
    - principal component analysis
- semi-supervised learning
- reinforcement learning
    - deep learning

>>>>

- supervised learning
    - continuous outcome (i.e., regression)
      - linear regression
      - lasso regression
      - ridge regression
      - elastic net regression
    - categorical outcome (i.e., classification)
      - logistic regression
      - support vector machine
      - random forest
      - extreme gradient boosting
- unsupervised learning
    - clustering
    - principal component analysis
- semi-supervised learning
- reinforcement learning
    - deep learning
- ensemble
